Exercise 2: Quality Attributes and Architectural Design of the HS07 System

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex

\begin{itemize}
\item \emph{Performance.} HS07 should be performant so that a large
  number of thermometers and radiators may be part of the system.
\item \emph{Modifiability.} It must be possible to modify HS07 to
  include new types of sensors and actuators.
\item \emph{Usability.} It must be easy to use the gateway user interface.
  include new types of sensors and actuators.
\item \emph{Availability.} The system must be responsive and available for 
both users and hardware elements.
\item \emph{Safety.} The system must take care not to overheat the actuators.
\end{itemize}

@@

\begin{enumerate}
\item A sensor stops responding.
\item Someone leaves a window open in an ice storm, and the sensors makes the actuators overheat.
\item Someone accidentally puts their stiletto through the Ethernet switch, disabling networking 
parts of the system.
\item After an average user has used the system for less than one hour e or she has mastered the basic functionality
\item The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute.
\item The user logs on externally through the gateway and has access to the home control within 1 second of verifying 95\% of the time (5 seconds 4\% of the time, more than 5 seconds (unscheduled down-time) less than 1\%) is connected and has access
\item The user attempts to turn radiator above normal safety-level from outside the house and receive an error back explaining the infraction. The system stores a log of the episode and refuse to execute the command.
\item An attacker attempts to log on using an invalid password and is thrown off. The next attempt will take longer (progressive log-on time). The system stores a log of the failed attempt.
\item Developer updates and tests the code base with a new device type in 1 day.
\item A technical installer installs a system at a customer in no more than 4 hours.
\item Developer updates the code base to a new requirement in less than 30 man-hours
\item Configuration manager installs an upgrade remotely at a customer in no more than 30min
\item Tester/debugger extract debugging information directly from running system at customer.
\item The system handles 50 sensors and 50 actuators.
\item The system handles up to 50 rooms/groups with one or more sensors and actuators in each group.
\item The user changes the preferred temperature from a web page and the system adapts to it.
\item The customer requests that the system runs on other hardware (e.g. less power consuming hardware).
\item The customer requests that the system interfaces with other sensors/actuators?
\item The developer modifies the sensor poll interval at build time.
\item The system turns off all actuators when sensors cannot be probed.
\item The end user adds a sensor/actuator at runtime which registers automatically.
\end{enumerate}

@@

% =====================================================================
\section{Evaluation}
% =====================================================================
\section{Architectural Design}
% =====================================================================
\bibliography{paper}
\bibliographystyle{apalike}

>>>>

\begin{itemize}
\item \emph{Performance.} HS07 should be performant so that a large
  number of thermometers and radiators may be part of the system.
\item \emph{Modifiability.} It must be possible to modify HS07 to
  include new types of sensors and actuators.
\item \emph{Usability.} It must be easy to use the gateway user interface.
  include new types of sensors and actuators.
\item \emph{Availability.} The system must be responsive and available for 
both users and hardware elements.
\end{itemize}

@@

\begin{enumerate}
\item A sensor stops responding.
\item Someone leaves a window open in an ice storm, and the sensors makes the actuators overheat.
\item Someone accidentally puts their stiletto through the Ethernet switch, disabling networking 
parts of the system.
\item After an average user has used the system for less than one hour e or she has mastered the basic functionality
\item The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute.
\item The user logs on externally through the gateway and has access to the home control within 1 second of verifying 95\% of the time (5 seconds 4\% of the time, more than 5 seconds (unscheduled down-time) less than 1\%) is connected and has access
\item The user attempts to turn radiator above normal safety-level from outside the house and receive an error back explaining the infraction. The system stores a log of the episode and refuse to execute the command.
\item An attacker attempts to log on using an invalid password and is thrown off. The next attempt will take longer (progressive log-on time). The system stores a log of the failed attempt.
\item Developer updates and tests the code base with a new device type in 1 week.
\item A technical installer installs a system at a customer in no more than 4 hours.
\item Developer updates the code base to a new requirement in less than 30 man-hours
\item Configuration manager installs an upgrade remotely at a customer in no more than 30min
\item Tester/debugger extract debugging information directly from running system at customer.
\item The system handles 50 sensors and 50 actuators.
\item The system handles up to 50 rooms/groups with one or more sensors and actuators in each group.
\item The user changes the preferred temperature from a web page and the system adapts to it.
\item The customer requests that the system runs on other hardware (e.g. less power consuming hardware).
\item The customer requests that the system interfaces with other sensors/actuators?
\item The developer modifies the sensor poll interval at build time.
\item The system turns off all actuators when sensors cannot be probed.
\item The end user adds a sensor/actuator at runtime which registers automatically.
\end{enumerate}

@@

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability  \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The developer \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Whishes to add a new device type \\
  \cline{2-3}
  & {\bfseries Artifact} & The code \\
  \cline{2-3}
  & {\bfseries Environment:} & At design time \\
  \cline{2-3}
  & {\bfseries Response:} & Modification is made with no side effects (all unit-/integrationtest pass)\\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Modification is made and tested within 1 week \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor? Is it necessary to make structual modifications to the user interface? Does the change have architectural impact? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#1 - QA Scenario 9/11/18}
\end{center}
\end{table}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Attaches a new actuator \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & Maintanance operation \\
  \cline{2-3}
  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Within one minute \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#2 - QA Scenario 5/21}
\end{center}
\end{table}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Remotely change the temperature \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & Normal operation \\
  \cline{2-3}
  & {\bfseries Response:} & Begin adjusting to temperature \\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When the command is received by the radiator? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#3 - QA Scenario 6/16}
\end{center}
\end{table}

% =====================================================================
\section{Evaluation}
The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined scenarios
is discussed in the following chapters.
\subsection{Refined Scenario \#1}
The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.

\subsection{Adding new sensor/actuator}
Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.

\subsection{Adding different types}
Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.

We have chosen to focus in the integration of new types of thermometers and actuators.

\subsection{Refined Scenario \#2}
Because of the Client/Server architecture of the gateway, adding new actuators is currently possible without architectural changes. With the current design, only actuators of the radiator type can be added.

\subsection{Refined Scenario \#3}
In order to have the system encompass a user interface that allows a user to change the preferred temperature from a web page, the gateway subsystem needs to become the Model part of a MVC pattern.

This functionality is not part of the current architecture, even though it supports commands on the gateway. At least one new command needs to be implemented on the gateway.

% =====================================================================
\section{Architectural Design}

\subsection{Refined Scenario \#1}
\subsection{Tactics}
\subsection{Modifiablity}
\subsection{Testability}
\subsection{Refined Scenario \#3}
\subsection{Tactics}
\subsection{Availability}
We will use a passive redundancy system for the http service in order to handle the response measurement. 
We consider the risk of hardware failure to small to use hardware redundancy, given the expected low amount of traffic on the system.
\subsection{Usability}
The web interface must...

% =====================================================================
\bibliography{paper}
\bibliographystyle{apalike}
